---
RITA Flight Delays by Wenny Wu
========================================================

Load the data:

```{r echo=FALSE, message=FALSE, warning=FALSE}
setwd("C:/Users/Administrator/Desktop/Udacity_Data_Analyst_Nanodegree/Udacity-DAND-P6_Flights_Visualization/data")

# Load the Data
df <- read.csv('data.csv')

library(ggplot2)
library(dplyr)
library(gridExtra)
```

Let's look at the data and summary statistics:

```{r echo=FALSE, message=FALSE, warning=FALSE}
str(df)
names(df)
summary(df)
```

Data cleaning - re-format the date:

```{r echo=FALSE, message=FALSE, warning=FALSE}
df$date <- as.Date(paste(df$year, df$X.month, 1, sep='-'), format="%Y-%m-%d")
summary(df$date)
```

Create a table of delay information:

```{r echo=FALSE, message=FALSE, warning=FALSE}
fl <- df %>%
  group_by(date, year, airport, airport_name) %>%
  summarize(arrivals = sum(arr_flights),
            delayed = sum(arr_del15),
            carrier_delays = sum(carrier_ct),
            weather_delays = sum(X.weather_ct),
            nas_delays = sum(nas_ct),
            security_delays = sum(security_ct),
            late_aircraft_delays = sum(late_aircraft_ct)) %>%
  transform(on_time = 1 - delayed/arrivals)

fl <- fl[complete.cases(fl),]
```

Plot delay information by airport:

```{r echo=FALSE, message=FALSE, warning=FALSE}
pf <- ggplot(data = fl, aes(x = date, y = on_time)) +
  geom_line(aes(color = airport))

pf + theme(legend.position="none")
```

Plots are too messy. Subset to look at only some airports:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# aggregate by airport
agg_apt <- fl %>%
  group_by(airport) %>%
  summarize(monthly_avg = mean(arrivals),
            arrivals = sum(arrivals))

big_airports <- subset(agg_apt, monthly_avg >= quantile(monthly_avg, 0.98))$airport
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
fb <- subset(df, is.element(airport, big_airports)) %>%
  group_by(year, airport, airport_name) %>%
  summarize(arrivals = sum(arr_flights),
            delayed = sum(arr_del15),
            carrier_delays = sum(carrier_ct),
            weather_delays = sum(X.weather_ct),
            nas_delays = sum(nas_ct),
            security_delays = sum(security_ct),
            late_aircraft_delays = sum(late_aircraft_ct)) %>%
  transform(on_time = 1 - delayed/arrivals)

fb <- fb[complete.cases(fb),]
```

Plot delay information by largest airports:

```{r echo=FALSE, message=FALSE, warning=FALSE}
pa1 <- ggplot(data = fb,
       aes(x = year, y = on_time)) +
  geom_line(aes(color = airport_name)) +
  scale_x_continuous(limits=c(2003, 2015), breaks=c(2003:2015))
  
pa1
```

Plot more detailed delay info by airport - carrier delays, weather delays, nas delays, security delays, and late aircraft delays:

```{r echo=FALSE, message=FALSE, warning=FALSE}
pc <- ggplot(data = fb,
       aes(carrier_delays, fill = airport)) +  
  geom_histogram(binwidth = 1000) + 
  ggtitle('Carrier Delays')

pw <- ggplot(data = fb,
       aes(weather_delays, fill = airport)) +  
  geom_histogram(binwidth = 500) + 
  ggtitle('Weather Delays')

pn <- ggplot(data = fb,
       aes(nas_delays, fill = airport)) +  
  geom_histogram(binwidth = 5000) + 
  ggtitle('National Air System Delays')

ps <- ggplot(data = fb,
       aes(security_delays, fill = airport)) +  
  geom_histogram(binwidth = 50) + 
  ggtitle('Security Delays')

pl <- ggplot(data = fb,
       aes(late_aircraft_delays, fill = airport)) +  
  geom_histogram(binwidth = 5000) + 
  ggtitle('Late Aircraft Delays')

grid.arrange(pc, pw, pn, ps, pl, ncol=3)
```

Write to csv:

```{r}
write.csv(fb, file="data_airport.csv", row.names=FALSE)
```
